---
title: "Nordic ELISA Analysis"
output: html_notebook
---

Load in libraries.
```{r}
library(tidyverse)
library(reshape)
library(ggpubr)
library(rstatix)
library(ggsignif)
```

Read in data and drop the unwanted analytes.
```{r}
orig_data <- read.csv('/~/Data/NordicData_reformatted.csv')
orig_data$C7M[which(orig_data$C7M=='ND')] <- NA
data <- orig_data %>% select('Tube', 'Cohort', 'Gender', 'Age', 'C1M_HP', 'C3M_HP', 'PRO.C5_HP', 'PRO.C7_HP', 'CTX.III_HP')
data
```

Look at data normality.
```{r}
ggqqplot(data, x = 'C1M_HP', facet.by = 'Cohort')
ggqqplot(data, x = 'C3M_HP', facet.by = 'Cohort')
ggqqplot(data, x = 'PRO.C5_HP', facet.by = 'Cohort')
ggqqplot(data, x = 'PRO.C7_HP', facet.by = 'Cohort')
```

Perform comparisons for designated analytes.
```{r}
columns <- c("C1M_HP", "C3M_HP", "PRO.C5_HP", "PRO.C7_HP", "CTX.III_HP")

p_values <- sapply(columns, function(col) {
  t.test(as.formula(paste(col, "~ Cohort")), data = data)$p.value
}, USE.NAMES = TRUE)

p_values
```

Define functions to calculate standard error and fold change.
```{r}
# Function to calculate standard error
calculate_se <- function(data, cohort, column) {
  cohort_data <- data %>%
    na.exclude() %>%
    filter(Cohort == cohort) %>%
    pull(!!sym(column))
  
  se <- sd(cohort_data) / sqrt(length(cohort_data))
  return(se)
}

# Function to calculate fold change (FC)
calculate_fc <- function(data, control_column, cohort_column) {
  control_sum <- data %>%
    na.exclude() %>%
    filter(Cohort == 'Control') %>%
    summarise(sum_value = sum(!!sym(control_column))) %>%
    pull(sum_value)
  
  dup_sum <- data %>%
    na.exclude() %>%
    filter(Cohort == 'Dup') %>%
    summarise(sum_value = sum(!!sym(cohort_column))) %>%
    pull(sum_value)
  
  fc <- dup_sum / control_sum
  return(fc)
}
```

Calculate SE.
```{r}
se_results <- map(columns, function(col) {
  list(
    Control = calculate_se(orig_data, 'Control', col),
    Dup = calculate_se(orig_data, 'Dup', col)
  )
})

names(se_results) <- columns

data.frame(se_results)
```

Calculate FC
```{r}
fc_results <- map(columns, function(col) {
  calculate_fc(orig_data, col, col)
})

names(fc_results) <- columns

data.frame(fc_results)
```

Calculate effect size.
```{r}
cd_df <- data.frame()

for (analyte in columns) {
  # Create a dynamic formula using as.formula
  formula <- as.formula(paste(analyte, "~ Cohort"))
  
  res <- cohens_d(formula, data = data, var.equal = TRUE)
  
  cd_df <- bind_rows(cd_df, res)
}

cd_df
```

Define plotting function.
```{r}
violin_plot <- function(data, analyte) {
  
  df <- ggplot(data, aes(x = Cohort, y = !!sym(analyte))) +
    geom_violin(width = 0.5, aes(color = Cohort)) +
    geom_boxplot(outlier.shape = NA, width = 0.2, coef = 0, aes(color = Cohort)) +
    geom_signif(comparisons = list(c("Control", "Dup")), map_signif_level = TRUE, textsize = 8.25) +
    scale_color_manual(values = c("#008000", "#800080")) +
    geom_jitter(alpha = 0.8, width = 0.1, aes(color = Cohort))
  
  return(df)
}
```

Prepare plots (with some customization based on analyte).
```{r}
C1M <- violin_plot(data, 'C1M_HP') +
  coord_cartesian(ylim = c(80, 240)) +
  theme(legend.position="none", text = element_text(size = 18), axis.title.x=element_blank()) + 
  ylab("C1M (ng/mL)")
C1M

proC5 <- violin_plot(data, 'PRO.C5_HP') +
  coord_cartesian(ylim = c(850, 1705)) + 
  theme(legend.position="none", text = element_text(size = 18), axis.title.x=element_blank()) +
  ylab("Pro-C5 (ng/mL)")
proC5

proC7 <- violin_plot(data, 'PRO.C7_HP') +
  coord_cartesian(ylim = c(20, 85)) +
  theme(legend.position="none", text = element_text(size = 18), axis.title.x=element_blank())+ 
  ylab("Pro-C7 (ng/mL)")
proC7

# Prepped CTX plot slightly differently as the violin is so thin the box size needed to be adjusted
CTX <- ggplot(data, aes(x = Cohort, y = CTX.III_HP)) + geom_violin(width = 0.8, aes(color = Cohort)) +
  geom_boxplot(outlier.shape = NA, width = 0.1, coef = 0, aes(color = Cohort)) + 
  geom_signif(comparisons = list(c("Control","Dup")), map_signif_level=TRUE, textsize = 8.25) + 
  scale_color_manual(values = c("#008000", "#800080")) +  
  geom_jitter(alpha = 0.8, width = 0.1, aes(color = Cohort)) +
  theme(legend.position="none", text = element_text(size = 18), axis.title.x=element_blank()) +
  ylab("CTX-III (ng/mL)") + 
  coord_cartesian(ylim = c(20, 605))
CTX
```

Create a df of pro/mature collagen ratios.
```{r}
ratio_df <- bind_cols(orig_data[,1:4],
          "Collagen1" = orig_data$PRO.C1_HP/orig_data$C1M_HP,
          "Collagen3" = orig_data$PRO.C3_roHP/orig_data$C3M_HP,
          "Collagen4" = orig_data$PRO.C4_HP/orig_data$C4M_HP,
          "Collagen6" = orig_data$PRO.C6_roHP/orig_data$C6M_HP)
          # Excluding PRO.C7 as the majority of columns are effectively undetected
          #"Collagen7" = orig_data$PRO.C7_HP/orig_data$C7M)
ratio_df
```

Plot the collagen I ratio.
```{r}
col1ratio <- violin_plot(ratio_df, 'Collagen1') +
  coord_cartesian(ylim = c(min(ratio_df$Collagen1), max(ratio_df$Collagen1)+0.3)) +
  theme(legend.position="none", text = element_text(size = 18), axis.title.x=element_blank()) +
  ylab("Pro-C1/C1M")
col1ratio
```

Export and save all figures.
```{r}
ggsave(filename = "C1M.png", plot = C1M, path = "~/Desktop",  device="png", dpi=600)

ggsave(filename = "Col1Ratio.png", plot = col1ratio, path = "~/Desktop",  device="png", dpi=600)

ggsave(filename = "proC5.png", plot = proC5, path = "~/Desktop",  device="png", dpi=600)

ggsave(filename = "proC7.png", plot = proC7, path = "~/Desktop",  device="png", dpi=600)

ggsave(filename = "CTXIII.png", plot = CTX, path = "~/Desktop",  device="png", dpi=600)

ggsave(filename = "Col1Ratio.png", plot = col1ratio, path = "~/Desktop",  device="png", dpi=600)

```
